---
title: "Prepare data for time-serie analysis"
output: html_document
---

``````{r}
rm(list=ls())
````````

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####preamble

Here we are going to prepare the data in order to analyse just evolution.

#### libraries
````{r}
library(ggplot2)
library(data.table)
library(zoo)
library(dplyr)
library(tidyr)
```````

#### data

````{r}
b<-fread("EEpapers_topics_byyear_05.csv", header=TRUE)
b<-as.data.frame(b)
length(unique(b$topic_label))
````
Keep the important columns
````````{r}
b<-b%>%
  select(originaltopic, haldtopic,topic_label, topic_use, topic_majortaxon, topic_discipline, year, proportion_present, sd_proportion_present)
````````
filter 1900-2010. In fact we are going to start in 1896 since later we are going to use 5 year rollmeans
``````````{r}
cutoffyear<-1960
b2<-b%>%
filter(year>=cutoffyear-4)%>%
filter(year<=2010)
```````````````
number of topics:
````{r}
unique(b2$topic_label)
length(unique(b2$originaltopic))
`````
first we get the rolling mean.  
````````{r}

b2<-b2%>% group_by(topic_label) %>% 
mutate(rollmean = rollmean(x = proportion_present, 5, align = "right", fill = NA))%>%
filter(year>=cutoffyear)%>%
as.data.frame()
``````
 
let's first pull out all years in which proportion is zero. and then the maximum year in which each topic is zero

```````{r}
b3<-b2%>%
filter(proportion_present==0)%>%
group_by(topic_label)%>%
filter(year==max(year))%>%
select(topic_label, year)%>%
arrange(desc(year))%>%
rename(last_zero_year=year)%>%
as.data.frame()
````````````
Merge this dataframe with our original data. And filter all years after the last year zero date

```````{r}
b4<-merge(b2,b3, by="topic_label", all.x=TRUE)
b5<-b4%>%
  mutate(last_zero_year=ifelse(is.na(last_zero_year), cutoffyear, last_zero_year))%>%
  filter(year>last_zero_year)
```````
identify years in which rollmean >= 0.005 (0.5%à. And then filter for the first year in which rollmean >= 0.005
`````{r}
b6<-b5%>%
  filter(rollmean>=0.005)%>%
  group_by(topic_label)%>%
  filter(year==min(year))%>%
  select(topic_label, year)%>%
arrange(desc(year))%>%
rename(first_0.005_year=year)%>%
as.data.frame()
`````````
 
Let's merge this dataframe with our original data. And filter all years after first_0.005_year

``````{r}
b7<-merge(b5, b6, by="topic_label", all.x=TRUE)
b8<-b7%>%
  mutate(first_0.005_year=ifelse(is.na(first_0.005_year), cutoffyear, first_0.005_year))%>%
  filter(year>=first_0.005_year)
```````

Let's check if we have any zero values left:

`````````{r}
b8%>%
  filter(proportion_present==0)
```````


````````{r}
b9<-b8%>%
  group_by(topic_label)%>%
  summarize(N_years=length(year))%>%
  arrange(N_years)
b9
``````
One series is very short:  urban geography will have only 2 years since it begins in 2008 and our data run till 2010.  
Let's remove it. 

``````{r}
b10<-b8%>%
  filter(topic_label!="urban ecology")
```````

Now let's normalize our proportions by the first year
```````{r}
b11<-b10%>%
  select(topic_label, year, proportion_present)%>%
  group_by(topic_label)%>%
  filter(year==min(year))%>%
  rename(start_year=year, start_year_proportion_present=proportion_present)
````````
merge this with our data and then normalize the year and proportions
````````{r}
b12<-merge(b10, b11, by="topic_label")
b12<-b12%>%
  arrange(topic_label, year)%>%
  mutate(norm_year=year-start_year, norm_prop=proportion_present/start_year_proportion_present)
``````



```````{r}
b13<-b12%>%
  select(-c("last_zero_year","first_0.005_year" ))
write.csv(b13, "EE_proportions_1960_2010_for_TS.csv", row.names=FALSE)
```````

