---
title: "Finalscript"
author: "orianebasso, Armand Marie Leroi, Ben Lambert"
date: "01/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())

```
##### libraries
`````````{r}
setwd("~/evolutionsfinalscripts")

library(rstan)
library(tidyverse)
library(ggplot2)
library(data.table)
library(zoo)
library(dplyr)
library(tidyr)
rstan_options(auto_write = TRUE)


````````

##### data

``````{r}
b<-fread("EE_proportions_1960_2010_for_TS.csv", header=TRUE)
b<-as.data.frame(b)

````````

``````{r}
N <- length(unique(b$topic_label)) #number of series 
Ob<-length(b$topic_label) #number of observations

T<-b%>%
  group_by(topic_label)%>%
  summarize(N_years=length(year)) 
T<-T$N_years #length of series in a vector
          

index<-c(1)
subset<-c(2:169)
for (i in subset) {index[i]<- index[i-1]+T[i-1] }  #start position of each study

X <- as.vector(b$norm_prop) %>% log #vector of all observations


groups<-c() #identify the series to which each observation in X belongs
subset<-c(1:169)
for (i in subset) { 
  for (k in index[i]:(index[i]+T[i]-1)) {groups[k]<-i } }
  
b<-mutate(b,topicnumber=groups)

series_start<-rep(0,Ob)
subset2<-c(1:Ob)
for (i in 1:N){
  for (k in subset2) {if (k==index[i]) series_start[k]<-1
}
}


```````



##### equation 5
````````{r}
fit <- stan('equation_five.stan',
            data=list(X=X,Ob=Ob,N=N,index=index,groups=groups,series_start=series_start),
            iter=2000,chains=4,control=list(adapt_delta=0.95,stepsize=0.01))
fit
`````````

`````{r}
r<- rstan::extract(fit, 'rho')
r<-as.data.frame(r)
r1<-r%>%
  mutate(iteration=1:nrow(r))
dim(r1) 
```````

summarize the rho posteriors for each variable:
```````{r}
r2<-r1%>%
  pivot_longer(-c(iteration), names_to="topicnumber", values_to="estimate")%>%
  as.data.frame()%>%
  mutate(topicnumber=gsub("rho.", "", topicnumber))%>%
  group_by(topicnumber)%>%
  summarize(median=quantile(estimate, 0.5), LCI=quantile(estimate, 0.025), UCI=quantile(estimate, 0.975))%>%
  arrange(desc(median))%>%
  mutate(mode=ifelse(UCI>=1, "Random Walk", "Mean Reverting"))
````````

How many RW and MR series are there:

``````````{r}
r2%>%
group_by(mode)%>%
summarize(N=length(topicnumber))
r2<-rename(r2,median5=median,UCI5=UCI,LCI5=LCI)
write.csv(r2, "ourdata_rw_MR.csv", row.names = FALSE)
```````


##### equation 6


``````{r}
median6<-c()
LCI6<-c()
UCI6<-c()
subset<-c(1:169)
for (i in subset) { 
  print(i)
datastan2<-list(X=X[index[i]:(index[i]+T[i]-1)],L=T[i])
#print(data_stan2)
fit <- stan("rw_driftbis.stan",data=datastan2,iter=2000,chains=4)
#print(fit)
delta <- rstan::extract(fit, "aDrift")[[1]] # extract delta
delta1<-as.data.frame(delta)# extract delta
delta2<-delta1%>%# summarize delta
summarize(LCI6=quantile(delta, 0.025), median6=quantile(delta, 0.5), UCI6=quantile(delta,0.975))
print(delta2)
median6[i]<-delta2$median6
LCI6[i]<-delta2$LCI6
UCI6[i]<-delta2$UCI6
}
result6<-cbind(LCI6, median6, UCI6)
result6<-as.data.frame(result6)
topics<-unique(b$topicnumber)
result6$topicnumber<-topics 
topics2<-unique(b$topic_label)
result6$topiclabel<-topics2
write.csv(result6, "ourdata_rw_drift_estimates.csv", row.names = FALSE)

```````

##### equation 7

`````{r}
median7<-c()
LCI7<-c()
UCI7<-c()
subset<-c(1:169)
for (i in subset) { 
  print(i)
datastan2<-list(X=X[index[i]:(index[i]+T[i]-1)],L=T[i])
#print(data_stan2)
fit <- stan("stationarybis.stan",data=datastan2,iter=2000, chains=4)
#print(fit)
delta <- rstan::extract(fit, "aDrift")[[1]] # extract delta
delta1<-as.data.frame(delta)# extract delta
delta2<-delta1%>%# summarize delta
summarize(LCI7=quantile(delta, 0.025), median7=quantile(delta, 0.5), UCI7=quantile(delta,0.975))
print(delta2)
median7[i]<-delta2$median7
LCI7[i]<-delta2$LCI7
UCI7[i]<-delta2$UCI7
}
result<-cbind(LCI7, median7, UCI7)
result7<-as.data.frame(result)
topic<-unique(b$topicnumber)
result7$topicnumber<-topic 
topics3<-unique(b$topic_label)
result7$topiclabel<-topics3
write.csv(result7, "ourdata_mrp_drift_estimates.csv", row.names = FALSE)
````````

##### Classification
```{r}


RW <- subset(r2,UCI5>=1) %>%
merge(result6,by="topicnumber") %>%
mutate(mode=ifelse((LCI6<=0)&(UCI6>=0), "URW", "BRW"))

MR <- subset(r2,UCI5<1) %>%
merge(result7,by="topicnumber") %>%
mutate(mode=ifelse((LCI7<=0)&(UCI7>=0), "UMR", "BMR"))

nouveau<-bind_rows(RW,MR) 

group_by(nouveau,mode) %>%
  summarize(N=length(topicnumber))



```

`````````{r}
finalresult<-b[,c("topicnumber","topic_label")] %>% 
  merge(nouveau,by="topicnumber")
write.csv(finalresult, "ourdata_finalresult.csv", row.names = FALSE)

`````````

